Increasing Productivity Amid Stable Nutrient Regimes in Rhode Island Lakes and Reservoirs

### 2019-05-14 manuscript/manuscript.Rmd
updated figure caption to italics a
fixed schematic all the way down

- Parameter by parameter discussion of trends
- How do trends break down by trophic state?
- Local scale vs Regional/Global scale
- RI is downstream - Bryan will write

@@

- Provides feedback on long term changes for managers and stakeholders
- Season Length Changes
- How to stabilize trends?  More reductions in nutrients?  Engineered solutions to slow warming? Look to individual lakes that buck the trend for suggestions (e.g. local landcover may provide a buffer)

@@

- Why this is possible - Fantastic WW data
- Use of site specific z-scores
    - Other approaches available, but this one is useful becuase it provided readily interpretable results, robust to variations in sampling stations (i.e. added later in period, removed, etc.), and based off of approach with long history of use in exploring long term trends [@hansen2006global, hansen2010global, @jones1996calculating, @jones1999surface]

>>>>

As previously mentioned, both temperature and chlorophll *a* show increasing trends from 1993 to 2016 in Rhode Island lakes and reservoirs, while total nutrients and the nitrogen:phosphorus ratio are all relatively stable.  These stable nutrient regimes may be partly explained by

- Parameter by parameter discussion of trends
- Local scale vs Regional/Global scale
    - Why diffs?  
        - LULC?
        - Different trends
- Other trends?
    - monthly/seasonal
    - How do trends break down by trophic state?

@@

- Provides feedback on long term changes for managers and stakeholders
- Season Length Changes
- How to stabilize trends?  
    - More reductions in nutrients?  
    - Engineered solutions to slow warming? 
    - Look to individual lakes that buck the trend for suggestions (e.g. local landcover may provide a buffer)
    - stable nutrients doesn't mean they are low, just not getting worse
    - interventions to reduce nutrients could offset the suggested increase caused by temperature
        - e.g. freshwater mussels restoration

@@

- Why this is possible - Fantastic WW data
- Use of site specific z-scores
    - Other approaches available, but this one is useful because it provided readily interpretable results, robust to variations in sampling stations (i.e. added later in period, removed, etc.), and based off of approach with long history of use in exploring long term trends [@hansen2006global, hansen2010global, @jones1996calculating, @jones1999surface]
